---
title: "Group Homework 1"
author: "STAT011-S23"
date: 'Due: X/X/23'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(nycflights13)
```



Make sure the R packages `tidyverse` and `nycflights13` are installed on your version of R. Call these packages into your library with the commands
```{r, eval=FALSE}
library("tidyverse")
library("nycflights13")
```
In this package is a data set called `flights` which records information for different flights. A table (which is actually called a ``tibble`` in the R `tidyverse`) called `airlines` defines the abbreviations for the different airline carriers in the `flights` data set. To see these objects type in the commands
```{r, eval=FALSE}
flights
airlines
```
Choose one of the 16 different carriers to study in this homework assignment. Then run the code below to define a subset of the `airlines` data set for any one of the carriers you select. You will use this subset of the data for Problems 1-5 and it will be referred to as `myData`.
```{r, eval=FALSE}
myData <- flights %>% na.omit() %>% filter(carrier=="MQ")  
dim(myData)
## you can change AA to whichever of the 16 codes you'd like from the "airlines" tibble
write.table(myData, "Data/EnvoyAir_flights.txt", sep = ",", row.names=FALSE)
```


## Problem 1 

Answer the following questions about you data set called `myData`:

  1. What constitutes an observational unit? 
  
  2. What are the different variables being collected?
  
  3. Which of the variables are quantitative and which are categorical? 

Hint: You can use the following code to extract the column names of the data set `flights`.
```{r, eval=FALSE}
names(flights)
```


# Problem 2

Referencing the R notes from class, for the variable `distance` perform a one variable exploratory data analysis by providing:

  1. A labeled histogram __with a reasonable number of bins__;

  2. A labeled box plot;

  3. A five-number summary.


# Problem 3

Referencing the R notes from class, for the variable `origin` perform a one variable exploratory data analysis by providing:

  1. A labeled bar chart;
  
  2. A table of frequencies of different airports of origin. 


# Problem 4

Using the code below, generate a scatter plot of the variables `dep_delay` and `arr_delay`. Write 1-2 sentences describing any apparent relationship between these variables and providing an explanation for this relationship. 

```{r, eval=FALSE}
ggplot(myData) + 
  geom_point(aes(x = dep_delay, y = arr_delay)) +
  labs(title="Scatter plot of flight data", x="Departure delay", y="Arrival delay")
```

# Problem 5

Does the variable `distance` look like it could be from a Normal (Gaussian) population? Justify your answer with a histogram and with 1-2 sentences. 


# Problem 6

How much of your course grade is determined by your participation in class? List three different ways to get participation credit. 


# Note to future self: 

Remove problem 3, have problem 5 be a part of problem 2, add more problems on mean and standard deviation concepts 

