---
title: "Quiz 2"
author: "STAT011-S20"
date: 'Due: 4/14/2020 midnight EDT on Gradescope'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(nycflights13)
```

**Disclaimer:** The data we are examining in this quiz is real and reflects the reality of our lives right now. I understand that this is a distressing topic to study. The reasons I chose this data to analyze in this quiz are the following: 

1. I want you to feel like the attention that you are giving to this class has a serious purpose and value; 

2. I want you to realize how powerful the methods that you're learning are in real life; 

3. I want you to be empowered to critically analyze the world around you. 

***

The World Health Organization reported the proportion of deaths from COVID-19 cases overall (globally) to be $0.04$. In the US so far, we have observed $20486$ deaths out of $525704$ cases of COVID-19. 

```{r}
prop.test(20486, n=525704, conf.level=0.95)
```

## Problem 1 (4 points)

(a) Based on the information above, plot on the line below, the lower and upper bounds of a $95\%$ confidence interval for the probability of death from COVID-19 in the US. Along this line also label the sample estimate, $\hat{p}$, and the margin of error. Finally, add to this line a point identifying the estimate from the World Health Organization. 

\vspace{1cm}

***

\vspace{1cm}

(b) Identify the population of interest for the confidence interval in part (a) and interpret the population parameter $p$.

\vspace{1cm}

(c) Check the conditions necessary to apply the Central Limit Theorem to derive the above confidence interval. Is there any reason to suspect that these conditions are not met? 

\vspace{1cm}

## Problem 2 (6 points)

As we all know, the impact of the COVID-19 virus is devastating across the world. We have begun to see however, that there are terrifying differences in the fatality rates according to race in the US.

According to the Bureau of Labor and Statistics, the racial and ethnic composition of the US roughly follows the table below. 

| Race/Ethnicity | Proportion of the population |
| ---- | ---- |
| Black | 0.134 |
| Hispanic or Latinx | 0.183 |
| White | 0.604 |
| Other | 0.079 |

\vspace{1cm}

(a) If the proportions in the table above apply to the $20486$ recorded COVID-19 deaths, approximately how many fatalities would we expect to see within each demographic group? 

\vspace{1cm}

(b) Produce four $95\%$ confidence intervals for the proportion of fatalities we'd expect to see across each of the four demographic categories based on the total number of fatalities, $n=20486$.


\vspace{1cm}


(c) Thus far, there have been $5748$ recorded deaths from COVID-19 that included information on the deceased's race or ethnicity. The observed proportion of these $5748$ fatalities across each group are shown below. 

| Race/Ethnicity | Proportion of deaths | 
| --- | --- | 
| Black | 0.34 |  
| Latinx | 0.16 |  
| White | 0.42 |  

Compare the observed proportions of fatalities by race in the table above with your confidence intervals estimating the proportion of fatalities by race from part (b). What do you observe? 

\pagebreak 

## Extra credit (2 points)

Reach out to someone you haven't spoken to recently (a classmate, an old friend, a relative, etc). You don't have to provide evidence of reaching out, just give me a brief description of who you reached out to and how you contacted them (like their relationship to you and their initials is sufficient). 

\vspace{1cm}

## Resources 

[1] https://www.apmresearchlab.org/covid/deaths-by-race

[2] https://www.bbc.com/news/health-51674743

[3] https://www.cdc.gov/coronavirus/2019-ncov/cases-updates/cases-in-us.html#2019coronavirus-summary

[4] https://www.census.gov/quickfacts/fact/table/US/PST045218

Note: All data used in this exam was current as of April 13, 2020.  

