---
title: "Exploring data"
subtitle: "Relationships between variables and the Normal model"
author: "STAT 011 with Prof Suzy"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
  #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
  self_contained: false # if true, fonts will be stored locally
#    includes:
#     in_header: "assets/mathjax-equation-numbers.html"
nature:
  #      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
navigation:
  scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---
  

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')


# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE, # display code as typed
  size = "small" # slightly smaller font for code
)

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library(tidyverse)
#library('gridExtra')
#library(broom)
data <- read_csv("Data/global_wealth_and_health.csv")
data2 <- data  %>% filter(!is.na(Value)) %>% 
               filter(Time=="2018") %>% 
               select(-c('Flag Codes', Flags, LOCATION, TIME, Time, DEMO_IND)) 
data3 <- pivot_wider(data2, id_cols=Country, names_from = Indicator, values_from = Value)
global_wealth_and_health <- data3[,-c(4,9,13)]
colnames(global_wealth_and_health) = c("Country", "FertRate", "LifeExpect", "MortRate", 
      "HIVPrev", "RuralPop", "GDP", "TotalExpenditure", "TotalPop", "SrPop", "YoungPop", "PovertyCount")
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```
     
## Today 

1.  Announcements 

2.  Hands on with RMarkdown 

3.  Group concept questions 

4.  Class discussion  

5.  Looking ahead 


---
## Announcements 

It is really important to read and follow all instructions! Also, all material you hand in for this class should be written in compplete sentences with proper punctuation. As a rule of thumb, pretend your work will be read by my grandfather who knows nothing about computers or R or statistics. 


In general, no late homework will be accepted. Each problem is worth 1 point and only graded as 0, 1/2, or 1 point credit per problem.  Recall you all have one drop HW grade!

Special accomodations for HW 1 

- For students who have been in class since day 1: you may submit corrections to HW 1 on Gradescope using the re-grade request option. 

- For students who have joined us after day 1: both HW 1 and 2 are both due by the beginning of class next Tuesday and you will not be able to submit corrections for either.  


---
## Announcements 

What should you do if you can't get your R Markdown file to knit? 


--
Email me your .Rmd file before the deadline. In the future there will be a penalty for submitting .Rmd files instead of PDFs uploaded through Gradescope. Anything emailed to me after the deadline will recieve a grade of 0.  



**Note:** More stats help is available starting today! Check the calendar for my explanded office hours and expect an announcement regarding new Muse sessions next week. 


---
## Hands on with RMarkdown 
### Preparation 

**Do this now:** 

- Open our course Moodle page.  

  * If you have a working version of R and RStudio on your laptop, open it.

  * If you are using the Swarthmore server, log in to rstudio.swarthmore.edu now. 

- Read over the R user guide uploaded to Moodle. 

---
## Hands on with RMarkdown 
### Overview 

By yourself, you will complete the following four steps today. We will spend $25$ mins on this activity. 

Step 1: Open a new R Markdown document, give the file a name and select the option to create an HTML file. Save this file with the name "testingR"

Step 2: Knit the document as-is and read through the RMarkdown code and output. 

Step 3: Complete the following tasks within the template the R automatically provided you for your RMarkdown document

Step 4: Submit your edited RMarkdown file (ends in .rmd) to me on Moodle. If you finish early, begin to work on HW 2! 

---
## Hands on with RMarkdown 
### Instructions

Complete the following tasks within the template the R automatically provided you for your RMarkdown document

  * Add a new section header using the `##` prefix followed by a space and then the title "Venting about R"
  
  * Spend a minute or two expressing your R fustrations and honest opinions about the coding part of this class. (Don't write more than 5 sentences... otherwise you may not finish the activity!)
  
---
## Hands on with RMarkdown 
### Instructions continued...  
  
  * Add another header (again using the `##` prefix followed by a space) with the title "Anyway, let's try R out..." 

  * In this section add a new R code chunk labeled "testplot" by typing the following exactly but without the #'s
  
```{r}`r''`
# {r testplot}
#     library(tidyverse)
#     head(cars)
```

    
---
## Hands on with RMarkdown 
### Instructions continued...
    
  * Use the code below to create a scatterplot for the two quantitative variables within the R code chunk "testplot"
  
```{r eval=FALSE}
  cars %>% ggplot(aes(x=speed, y=dist)) + geom_point()
```
  
  
   * Now create a new R code chunk labeled "testcalc" by typing the following but without the #'s which calculates a correlation matrix for the quantitative variables `speed` and `dist` 
  
```{r}`r''`
# {r testcalc}    
# cars %>% select(speed, dist) %>% cor

```
  
Write a sentence below this R code chunk interpreting the correlation between speed and stopping distance and then knit your document to see the output! 
  


---
## Group concept questions 
### 25 mins

Download or open the Class 4 R Markdown worksheet in RStudio. Make sure you can knit the template first! 

Then work with your group members to answer the $7$ questions. 

The group reporter will email me your group's completed RMarkdown worksheet for participation credit before the end of class. 

Raise your hand if your group has a question! 


---
## Class discussion and overview 


Relationships between variables 

   * two quantitative 
   
   * one quantitative, one categorical 
   
   * two categorical 
   
Determining if a quantitative variable seems to follow a Normal distribution 

  * Look for symmetry and unimodality 
  
  * Determine if the 68, 95, 99.7 percent rule applies 
  
  * Create a Normal Probability Plot and compare deviations from the diagonal line 


