---
title: "Intro to Categorical Data Analysis"
author: "STAT 011 with Prof Suzy"
institute: "Swarthmore College"
output:
  #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
  xaringan::moon_reader:
  self_contained: false # if true, fonts will be stored locally
#    includes:
#     in_header: "assets/mathjax-equation-numbers.html"
nature:
  #      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
navigation:
  scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
header-includes:
  - usepackage{xcolor}
---
  

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')


# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE, # display code as typed
  size = "small" # slightly smaller font for code
)

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library(tidyverse)
#library('gridExtra')
#library(broom)
data <- read_csv("Data/global_wealth_and_health.csv")
data2 <- data  %>% filter(!is.na(Value)) %>% 
               filter(Time=="2018") %>% 
               select(-c('Flag Codes', Flags, LOCATION, TIME, Time, DEMO_IND)) 
data3 <- pivot_wider(data2, id_cols=Country, names_from = Indicator, values_from = Value)
global_wealth_and_health <- data3[,-c(4,9,13)]
colnames(global_wealth_and_health) = c("Country", "FertRate", "LifeExpect", "MortRate", 
      "HIVPrev", "RuralPop", "GDP", "TotalExpenditure", "TotalPop", "SrPop", "YoungPop", "PovertyCount")
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 60%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```
     
## Today 

1. Details for inference concerning categorical variables

2. R coding practice - group worksheet 

3. Video and reflection prompt

4. Looking ahead  
  
  
---
## Key concepts 
### Statistical inference 

* The sampling distribution of the sample mean (or sample proportion) is a hypothetical distribution based on the idea of repeatedly obtaining a random sample from the population of interest. 


--
* Statistical inference requires us to have a larger population in mind.  


--
* The scope of the conclusions always depends on the quality of the data and the design of the study. Garbage in means garbage out, from the most basic to the most advanced statistical methods.   


---
## Inference for categorical variables  

Today we are going to discuss how to conduct the different Chi-square procedures in R and practice determining which procedure to use for different scenarios. 


The coding part for these chi-square procedures can be a little confusing, which is why it is so important to keep in mind what type of and how many variables are involved for each procedure. The complication comes from the way categorical data is represented in a data set. 


---
## Inference for categorical variables  

From a statisticians point of view... 


```{r echo=FALSE, eval=FALSE}
#table1 <- rbind(c(" ", "First variable", "Second variable", ". . . ", "Last variable"), c("First data point", " ", " ", " ", " "), c(".", " ", " ", " ", " "), c(".", " ", " ", " ", " "), c(".", " ", " ", " ", " "), c("Last data point", " ", " ", " ", " "))
#knitr::kable( table1 )
```

|    Observational Units    |  First variable |   Second variable   |    . . .    |    Last variable    |
|:------------|-----:|--------------:|--------------:|-------:|
|  First data point      |   |          |         |   |  
|  Second data point  |    |           |        |        |     
|  .   |     |        |           |        |     
|  .   |     |           |           |        |                
|  .   |     |          |           |        |             
|  Last $(n^{th})$ data point   |     |           |            |    . . .    |          

---
## Inference for categorical variables  

So for two categorical variables (say, class year and discipline of major) collected for each observational unit, we would represent that data set in the following way. 

|    Observational Units    |  Class year |   Major discipline   |   
|:------------|-----:|--------------:| 
|  Student 1    | Freshman   |   Natural Sciences       |        
|  Student 2  |    Freshman     |  Social Sciences      |     
|  .   |     |        |      
|  .   |     |           |    
|  .   |     |          |    
|  .   |  Junior   |    Social Sciences      |    
|  Student n   | Senior    |   Humanities        |       

---
## Inference for categorical variables  

When presenting such data however, we typically summarize the values of categorical variables within a particular data set in a table of counts, called a *contingency table*. This way it's easier for us to tell how many levels there are for each variable and we get a sense of the distribution of the data across each level. 


--
For example, 

|     |     |   Major discipline   |        |     
|:------------|-----:|--------------:|--------------:| 
|   Class year       |  Natural Sciences |  Social Sciences       |  Humanities       |  
|  Freshman      | 5   |   10   |    3   |  
|  Sophomore   |  3  |    5     |     0     |      
|  Junior   | 2   |   0     |   0    |        |    
|  Senior   |  0  |     1 |     0   |     

---
## Producing a contingency table in R 

Consider the data set in the `forcats` package called `gss_cat`. This is a sample of data from the General Social survey [General Social survey](https://gssdataexplorer.norc.org/)

```{r}
library(forcats)
head(gss_cat) 
```


---
## Producing a contingency table in R 

If you have a data set that contains categorical variables, you can create a contingency table with the following code: 

```{r}
gss_cat %>% group_by(marital, relig) %>%  ## specify which categorical variables you want to tabulate
            summarise(counts = n())  %>%  ## this specifies that we are creating a feature of our data set called "counts" and this is computed by counting up the number of observations using the function n()
            spread(relig, counts)    ##specify which counts you want to span the columns, this step is not necessary if you are creating a table of counts for a single categorical variable 
```


---
## Producing a contingency table in R 


For more instructions on creating tables in R with the `tidyverse` package, see this [page](http://analyticswithr.com/contingencytables.html).


---
## Producing a contingency table in R 
### Group work 

With your group, complete the Class 21 Worksheet R markdown template. (This can be found on Moodle or in our shared RStudio folder.)




---
## Inference for the distribution of a single categorical variable 
###Test of homogeneity

Are the different possible outcomes (that is, the, say $k$, different levels) of a (categorical) variable equally likely? 

\begin{cases}
H_0: p_1 = p_2 = p_3 = \dots = p_k \\
H_A: \text{At least one of the } k \text{ probabilities in the null}\\
\quad \quad \text{ is different from the others.} 
\end{cases}

--
```{r}
# chisq.test(x = vector_of_varb_counts)
```


--
**Q:** What happens if our categorical variable only has two levels? 

---
## Inference for the distribution of a single categorical variable 
###Goodness-of-fit test 

Does a hypothesized probability distribution for the (say, $k$) levels of your categorical variable fit your data well?

\begin{cases}
H_0: p_1 = p{1,0}, p_2 = p_{2,0}, p_3 = p_{3,0}, \dots p_k = p_{k,0} \\
H_A: \text{At least one of the } k \text{ probabilities stated in} \\
\quad \quad \text{ the null is incorrect.} 
\end{cases}


Just as in our other hypothesis tests, the values stated in the null $p_{1,0}, p_{2,0}, \dots$ are actual numbers (proportions/probabilities) between zero and one. 


--
```{r}
# chisq.test(x = vector_of_varb_counts, p = c(prob1_null, prob2_null, prob3_null, ..., probk_null))
```

---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence

Does the data suggest that one variable may be associated with another? 

\begin{cases}
H_0: \text{Categorical variable 1 is independent of categorical variable 2.} \\
H_A: \text{Categorical variable 1 and categorical variable 2 are not independent.} 
\end{cases}


Unlike every other test we have discussed in this class, there are no parameters involved in this test! 

We will go over the R code for this test next class. 
