---
title: "Stat 11 Week 9"
subtitle: "Review of Unit 2" 
author: "Prof. Suzy Thornton"
date: "Spring 2023"
lang: "en-US"
output:
  html_document:
    df_print: paged
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{xcolor}
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


<span style="color:purple">notecard</span>
<span style="color:blue">hypothetical</span>
<span style="color:green">new terms</span>


# Unit 2 topics 

* Collecting data

* Randomness and probability


# Week 4 - Collecting data from experiments and observational studies

* 

* 


## Example 




# Week 6 - From randomness to probability and the laws of probability

* 

* 


## Example 



# Week 7 - Random variables and probability models

* 

* 


## Example 


***

# Student Requested Examples 


## Student's t-distribution 

analogy for degrees of freedom parameter 

Model parameter(s): 

Sample estimate(s): 
[pic showing that 68/99/99/7 rule no longer applies]  


[T-table](https://www.sjsu.edu/faculty/gerstman/StatPrimer/t-table.pdf)

```{r}
myT.Quantile = 3
pt(myT.Quantile, df = 12)

myT.Probability = .25
qt(myT.Probability, df = 12)
```

### Quantiles and probabilities 

