---
title: "Collecting data for analysis"
author: "STAT 011 with Prof Suzy"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
  #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
  self_contained: false # if true, fonts will be stored locally
#    includes:
#     in_header: "assets/mathjax-equation-numbers.html"
nature:
  #      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
navigation:
  scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
---
  

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')


# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE, # display code as typed
  size = "small" # slightly smaller font for code
)

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library(tidyverse)
#library('gridExtra')
#library(broom)
#data <- read_csv("global_wealth_and_health.csv")
#data2 <- data  %>% filter(!is.na(Value)) %>% 
#               filter(Time=="2018") %>% 
#               select(-c('Flag Codes', Flags, LOCATION, TIME, Time, DEMO_IND)) 
#data3 <- pivot_wider(data2, id_cols=Country, names_from = Indicator, values_from = Value)
#global_wealth_and_health <- data3[,-c(4,9,13)]
#colnames(global_wealth_and_health) = c("Country", "FertRate", "LifeExpect", "MortRate", 
#      "HIVPrev", "RuralPop", "GDP", "TotalExpenditure", "TotalPop", "SrPop", "YoungPop", "PovertyCount")
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```
     
## Today 

1.  Homework 3 Problem 3 

2.  Lecturette on collecting data for statistical analysis  

3.  Reminders and announcements

4.  Reflection prompt and poll 


---
## Homework 3 Problem 3 

While only $5\%$ of babies have learned to walk by the age of $10$ months, $75\%$ are walking by $13$ months of age. Assuming the age at which babies develop the ability to walk can be described by a Normal model, find the mean and standard deviation of the distribution of the age at which babies learn to walk. 


Additional information: The lower $5^{th}$ quantile of a standard Normal distribution is $-1.65$ and the lower $75^{th}$ quantile is $0.67$. 

### 1. Strategize/outline Solution  (5 mins)

Note: If you've already solved this problem and are confident in your solution, try to think of a way to answer the question without the additional information. It turns out that this is actually possible to do with some extra creativity! 


---
## Homework 3 Problem 3 

While only $5\%$ of babies have learned to walk by the age of $10$ months, $75\%$ are walking by $13$ months of age. Assuming the age at which babies develop the ability to walk can be described by a Normal model, find the mean and standard deviation of the distribution of the age at which babies learn to walk. 


Additional information: The lower $5^{th}$ quantile of a standard Normal distribution is $-1.65$ and the lower $75^{th}$ quantile is $0.67$. 

### 2. Work with your group to outline a solution on a whiteboard. (10 mins)


---
## Homework 3 Problem 3 

While only $5\%$ of babies have learned to walk by the age of $10$ months, $75\%$ are walking by $13$ months of age. Assuming the age at which babies develop the ability to walk can be described by a Normal model, find the mean and standard deviation of the distribution of the age at which babies learn to walk. 


Additional information: The lower $5^{th}$ quantile of a standard Normal distribution is $-1.65$ and the lower $75^{th}$ quantile is $0.67$. 

### 3. A group reporter will start our discussion by presenting your group's strategy to the class. 


---
## Quick R Tip
### Finding normal distribution quantiles 

The following code will produce the lower $5^{th}$ quantile of a standard Normal distribution: 
```{r}
qnorm(p=0.05, mean=0, sd=1)
```


---
## Collecting data for statistical analysis 

- Samples vs populations 

  * Similarities: observational units and variables of interest  
  
  * Differences: parameters vs estimates, source of randomness, sizes  
  
- Getting a quality sample is of the utmost importance! 

  * Representative of the population 
  
  * As unbiased as possible 
  
- We can't always control the study type, but we can often control the way data is collected

  * Experiment 
  
  * Observational study 
  
- What is *significant*? 


---
## Reminders

* Community guidelines 

  - Scholarly community where we strive to make sure everyone is respected and heard. 
  
  - Use respectful language and not interrupt others. 
  
  - Assume good intentions. 
  
  - Ask questions. 
  
  - We will wear our masks throughout the entire class session. 

---
## Reflection prompt 
### Test prep 

Fill out the worksheet for today's class. 


## Prepare for Thursday's class 

Next class, we will spend out time on the following topics:

* Troubleshooting R tips and tricks 

* Critical thinking practice regarding collecting data for statistical analyses 


