---
title: "Statistical Methods in General"
author: "STAT 011 with Prof Suzy"
institute: "Swarthmore College"
output:
  #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
  xaringan::moon_reader:
  self_contained: false # if true, fonts will be stored locally
#    includes:
#     in_header: "assets/mathjax-equation-numbers.html"
nature:
  #      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
navigation:
  scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
header-includes:
  - usepackage{xcolor}
---
  

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')


# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE, # display code as typed
  size = "small" # slightly smaller font for code
)

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library(tidyverse)
#library('gridExtra')
#library(broom)
data <- read_csv("Data/global_wealth_and_health.csv")
data2 <- data  %>% filter(!is.na(Value)) %>% 
               filter(Time=="2018") %>% 
               select(-c('Flag Codes', Flags, LOCATION, TIME, Time, DEMO_IND)) 
data3 <- pivot_wider(data2, id_cols=Country, names_from = Indicator, values_from = Value)
global_wealth_and_health <- data3[,-c(4,9,13)]
colnames(global_wealth_and_health) = c("Country", "FertRate", "LifeExpect", "MortRate", 
      "HIVPrev", "RuralPop", "GDP", "TotalExpenditure", "TotalPop", "SrPop", "YoungPop", "PovertyCount")
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 60%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```
     
## Today 

1. Lecturette for inference concerning categorical variables (15 min)

2. Test 2 review and group work (20 min)

3. HW 8 coding focus (10 min)

4. Reflection prompt (10 min)

5. Looking ahead - due dates (5 min)

6. Looking ahead - final project and test 3 (10 min)


---
## Inference for the distribution of a single categorical variable 
###Test of homogeneity

Are the different possible outcomes (that is, the, say $k$, different levels) of a (categorical) variable equally likely? 

\begin{cases}
H_0: p_1 = p_2 = p_3 = \dots = p_k \\
H_A: \text{At least one of the } k \text{ probabilities in the null}\\
\quad \quad \text{ is different from the others.} 
\end{cases}

--
```{r}
# chisq.test(x = vector_of_varb_counts)
```



**Q:** What happens if our categorical variable only has two levels? 


--
**A:** In this case, the test for homogeneity is the same as the test for a single population proportion where the null is $H_0: p = 0.5$! 


---
## Inference for the distribution of a single categorical variable 
###Goodness-of-fit test 

Does a hypothesized probability distribution for the (say, $k$) levels of your categorical variable fit your data well?

\begin{cases}
H_0: p_1 = p{1,0}, p_2 = p_{2,0}, p_3 = p_{3,0}, \dots p_k = p_{k,0} \\
H_A: \text{At least one of the } k \text{ probabilities stated in} \\
\quad \quad \text{ the null is incorrect.} 
\end{cases}


Just as in our other hypothesis tests, the values stated in the null $p_{1,0}, p_{2,0}, \dots$ are actual numbers (proportions/probabilities) between zero and one. 


--
```{r}
# chisq.test(x = vector_of_varb_counts, p = c(prob1_null, prob2_null, prob3_null, ..., probk_null))
```

---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence

Does the data suggest that one variable may be associated with another? 

\begin{cases}
H_0: \text{Categorical variable 1 is independent of categorical variable 2.} \\
H_A: \text{Categorical variable 1 and categorical variable 2 are not independent.} 
\end{cases}


Unlike every other test we have discussed in this class, there are no parameters involved in this test! 


--
**Q:** What happens if both categorical variables only have two levels each? 


--
**A:** In this case, the test for independence is the same as the two sample test for a difference in proportions where the null is $H_0: p_1 - p_2 = 0$! 


---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence

Does the data suggest that one variable may be associated with another? 

\begin{cases}
H_0: \text{Categorical variable 1 is independent of categorical variable 2.} \\
H_A: \text{Categorical variable 1 and categorical variable 2 are not independent.} 
\end{cases}



```{r}
# chisq.test(x = two_way_contingency_table_as_a_matrix)
```

See the next few slides for details on the **three step coding process** for this test. 

---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence - Coding 

Recall the `gss_cat` data from last class. 
```{r}
library(forcats)
head(gss_cat) 
```
---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence - Coding 

**Step 1: Create a two-way contingency table of all possible combinations of levels** 

.scroll-small[
We can create a two-way contingency table, called `my_table` for two categorical variables (say, `marital` and `reilg`) with the following R code: 
```{r}
my_table <- gss_cat %>% filter(!is.na(marital) & !is.na(relig)) %>%  ## this eliminates any rows that are missing values for the two variables of interest 
                        group_by(marital, relig) %>%  ## specify which categorical variables you want to tabulate
                        summarise(counts = n())  %>%  ## this specifies that we are creating a summary called "counts" which counts the number of observation in each level of each variable
                        spread(relig, counts)         ## specify which variable you want in the columns 
my_table
```
]


---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence - Coding 

**Step 2: Convert your contingency table into a matrix.** 

.scroll-small[Then we can convert `my_table` into a matrix using the `as.matrix()` function. 
```{r}
my_matrix <- as.matrix(my_table)
my_matrix 
```
]


---
## Inference to understand the relationship between two categorical variables 
### Chi-squared test of independence - Coding 

**Step 3: Plug the matrix of counts into the `chisq.test()` function**
.scroll-small[
Now (*if we didn't have any empty cells*) we could plug this *matrix* into the `chisq.test()` function to do a test for the independence of the two categorical variables `marital` and `relig`. Since we do have at least one religious affiliation and marital status combination without any observations (e.g. "Hinduism" and "Widowed"), we cannot actually do a chi squared test for independence! (So the following code doesn't actually work but is meant to give you a sense of how to code this test in cases where it does apply.)
```{r}
#chisq.test(x = my_matrix)
```
]

