---
title: "Statistical inference for a proportion"
author: "STAT 011 with Prof Suzy"
institute: "Swarthmore College"
output:
  xaringan::moon_reader:
  #   css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
  self_contained: false # if true, fonts will be stored locally
#    includes:
#     in_header: "assets/mathjax-equation-numbers.html"
nature:
  #      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
navigation:
  scroll: false # disable slide transitions by scrolling
includes:
  in_header: mystyles.sty
header-includes:
   - \usepackage{amsmath}
---
  

```{r setup_pres, include=FALSE, echo=FALSE}
#devtools::install_github("ropenscilabs/icon")
#devtools::session_info('rmarkdown')


# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE, # display code as typed
  size = "small" # slightly smaller font for code
)

knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings

rm(list=ls())
library(tidyverse)
#library('gridExtra')
#library(broom)
data <- read_csv("Data/global_wealth_and_health.csv")
data2 <- data  %>% filter(!is.na(Value)) %>% 
               filter(Time=="2018") %>% 
               select(-c('Flag Codes', Flags, LOCATION, TIME, Time, DEMO_IND)) 
data3 <- pivot_wider(data2, id_cols=Country, names_from = Indicator, values_from = Value)
global_wealth_and_health <- data3[,-c(4,9,13)]
colnames(global_wealth_and_health) = c("Country", "FertRate", "LifeExpect", "MortRate", 
      "HIVPrev", "RuralPop", "GDP", "TotalExpenditure", "TotalPop", "SrPop", "YoungPop", "PovertyCount")
```

```{css, echo=FALSE}
pre {
  background: #FFBB33;
  max-width: 100%;
  overflow-x: scroll;
}

.scroll-output {
  height: 70%;
  overflow-y: scroll;
}

.scroll-small {
  height: 50%;
  overflow-y: scroll;
}
   
.red{color: #ce151e;}
.green{color: #26b421;}
.blue{color: #426EF0;}
```
     
## Today 

1. Inference for proportions in R 

2. Hypothesis tests for an unknown proportion  

3. Class 15 Worksheet 

4. CI for an unknown proportion 

5. Group work - sample size determination 

6. Looking ahead  



---
## R code
### Inference for a proportion 

To conduct an $\alpha$-level hypothesis test for an unknown proportion/probability:

```{r}
#prop.test(x=?, n=?, p=?, alternative = "?", conf.level = 1-alpha)
```


* $x =$ the observed number of successes 

* $n =$ the sample size (or number of trials) 

* $p$ is the unknown population proportion of successes (or the true probability of a success)


--
To find a $100 \times (1-\alpha) \%$ CI for an unknown proportion/probability:
```{r}
#prop.test(x=?, n=?, conf.level = 1-alpha)
```

--
<br>
Remember: confidence level $= 1 - \alpha$ and both are pre-set by the researcher (i.e. you!) 



---
### Hypothesis tests for a proportion 

The CLT states conditions under which $\hat{p} \sim N\left(p, \frac{p(1-p)}n\right).$


--
The test statistic associated with either a one-sided or two-sided hypothesis test for an unknown proportion, $p$ is 
$$\begin{eqnarray*}T.S. &=& \frac{\hat{p} - p_0}{\sqrt{\frac{p_0(1-p_0)}{n}}}\\ &=& \sqrt{\frac{n}{p_0(1-p_0)}}(\hat{p} - p_0) \\
&=& \sqrt{\frac{n}{p_0(1-p_0)}}\hat{p} - \sqrt{\frac{n}{p_0(1-p_0)}}p_0.\end{eqnarray*}$$

--
If we assume $H_0: p = p_0$ is correct, then
$$T.S. = \frac{\hat{p} - p_0}{\sqrt{\frac{p_0(1-p_0)}{n}}} \hspace{.5cm}\widetilde{\tiny H_{0}}  \hspace{.5cm} N(0,1).$$

---
### Hypothesis tests for a proportion - Group work 

Spend the next $7$ minutes working with your group to prove that 

$$E(T.S.) = 0 \text{ and } Var(T.S.) = 1$$



---
### Hypothesis tests for a proportion

1. A *p-value* is the conditional probability that the sampling distribution of $\hat{p}$ takes on the value you observe based on your data, $\hat{y}_{obs}$, or something more extreme, given that the null hypothesis is true. 

2. A *significance level* is a pre-specified value between $0$ and $1$ that indicates the researchers threshold for describing an observed value of $\hat{p}$ as rare (or, significant) under the null hypothesis. 


--
A p-value can be interpreted on it's own in the same way that a CI can be interpreted on it's own. However, if the purpose of your analysis is to make a decision, the p-value must be compared to your significance level. 


--
**Q:** What happens if we set $\alpha = 0$? 


--
**A:** This is essentially saying that you cannot be convinced that the data is rare or unusual. This will become more clear later in our discussion of confidence intervals. 


---
## Class 15 Worksheet 
### Problems 8 - 10 


**8.** $\hat{p}_{obs} = 18/30 = 0.6$ 

**9.** If $p=0.5$ and $\hat{p}_{obs} = 0.6$, then the observed sample proportion is $0.1$ units away from the mean of the sampling distribution. 


**10.** "What is the probability that, repeating this experiment again and again produces the same value of $\hat{p}$ or something larger" is mathematically equivalent to asking "What is the probability that, repeating this experiment again and again produces the same value of T.S. that you observe or something larger"? 


---
## Class 15 Worksheet - Group work

**Instructions:** Spend the next $7$ mins with your group going over what to fill in for the question marks in the code below. All lines of code should give you the same p-value. 

```{r}
# teststat_obs = ? 
#pnorm(teststat_obs, mean=0, sd=1, lower.tail=FALSE) 
## or 
# p_hat = ? 
#pnorm(p_hat, mean=0.5, sd=sqrt(0.5*(1-0.5)/30), lower.tail=FALSE) 
## or 
#prop.test(18, n=30, p=0.5, alternative= "greater", conf.level=?) 
```


---
### Confidence intervals for a proportion 

A $(1-\alpha)\times 100\%$ CI for an unknown proportion, $p$, is 
$$\hat{p}_{obs} \pm \left( z_{\alpha/2} \times \sqrt{\frac{\hat{p}_{obs}(1-\hat{p}_{obs})}{n}}\right),$$
where $z_{\alpha/2}$ is the lower $\alpha/2^{th}$ quantile of a $N(0,1)$ distribution. 


Both p-values and CIs use the same assumption that the CLT can be applied to the sampling distribution of the sample proportions. For a fixed level of significance (or confidence), both provide the same amount of specificity in determining how unusual (or rare) your observed data is. 

* P-values can help with making a decision 

* CIs can help estimate an *effect size*, that is, the actual size/magnitude of the unknown parameter 




---
### Confidence intervals for a proportion 

A $(1-\alpha)\times 100\%$ CI for an unknown proportion, $p$, is 
$$\hat{p}_{obs} \pm \left( z_{\alpha/2} \times \sqrt{\frac{\hat{p}_{obs}(1-\hat{p}_{obs})}{n}}\right),$$
where $z_{\alpha/2}$ is the lower $\alpha/2^{th}$ quantile of a $N(0,1)$ distribution. 


**New terminology:** 

1. $\sqrt{\frac{\hat{p}_{obs}(1-\hat{p}_{obs})}{n}}$ is called the *standard error* of the sampling proportion. This is almost the same thing as the standard deviation of the sampling distribution of the sample proportions but using $\hat{p}$ rather than $p$. 

2. Half of the width of the CI is $z_{\alpha/2} \times \sqrt{\frac{\hat{p}_{obs}(1-\hat{p}_{obs})}{n}}$, which is called the *margin of error*. 


--
**Q:** What happens if we set $\alpha =0$? 


--
**A:** Our $100\%$ CI for $p$ becomes the entire real number line: $(-\infty, \infty)$! 

---
## Group work 
### Sample size determination 

Consider the magician and his possibly biased (towards Heads) coin again. Suppose he proposes to let you determine how many times you want him to flip the coin before you decide to guess if the coin is biased or fair. If you guess correctly, the magician will pay you \$10. If you guess incorrectly, you must pay the magician \$10.

You decided that the margin of error you are willing to tolerate is $5\%$ and the significance level you wish to use is $\alpha = 0.02$. How many times are you going to ask the magician to flip the coin? 

<br>

**Instructions:** Work with your group to answer this question. When your group is finished, have the reporter write your answer on a piece of paper and hand it to me. When the time is up, we'll announce the correct answer. 


<br> 
--
**Hint:** If you don't have any previous data to use, how might you play it safe in choosing a value to substitute in for $\hat{p}$? 

