---
title: "In-Class Worksheet"
author: "STAT011 with Prof Suzy"
date: 'Week 11: Practicing with Hypothesis Tests'
output:
  html_document: default
  pdf_document: default
---
```{r setup, echo=FALSE, include=FALSE}
require(knitr)
library(tidyverse)
knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings
```

\vspace{1cm}
**Name:** _____________________________________

# Instructions:

There are five questions in this worksheet. You will work with your group members to answer each question. Before getting started, take a moment and reflect on ways in which you can show your assigned group members respect. You may also view this initial [list of examples](https://docs.google.com/document/d/1emP8SiS2IO-_blKdVmVp-CJTr0NLlKG4vzMnphH-P7E/edit?usp=sharing) that we will add to over the semester.

**Briefly,** in the space below, specify one way in which you will work to show your group members respect during today's lesson:


\vspace{1cm}


# Part 1: Proportions 

## 1. Calculating probabilities for a proportion 

The Centers for Disease Control and Prevention reports that $9.3\%$ of surveyed high school students reported in 2015 that they had smoked cigarettes in the past 30 days. A college has $522$ students in its freshman class. How likely is it that more than $10\%$ of them are smokers?


\vspace{2cm}

## 2.  Two-sided alternative Z-test for a proportion 

A national vital statistics report indicated that about $3\%$ of all births produced twins. Is the rate of twin births the same among very young mothers? Data from a large city hospital found that only $7$ sets of twins were born to $469$ teenage girls. 

a) Test an appropriate hypothesis and state your conclusion. Be sure the appropriate assumptions and conditions are satisfied before you proceed.

b) What would a Type I error mean in this situation? 

\pagebreak 

# Part 2ß: Means 

## 3. Calculating probabilities for a mean with the t-distribution 

Organizers of a fishing tournament believe that the lake holds a sizable population of largemouth bass. They assume that the weights of these fish have a model that is skewed to the right with a mean of $3.5$ pounds and a standard deviation of $2.2$ pounds.

a) Explain why a skewed model makes sense here.

b) Explain why you cannot determine the probability that a largemouth bass randomly selected (“caught”) from the lake weighs over $3$ pounds.

c) Each fisherman in the contest catches $5$ fish each day. Can you determine the probability that someone’s catch averages over $3$ pounds? Explain.

d) The $12$ fishermen competing each caught the limit of $5$ fish. What’s the probability that the total catch of $60$ fish averaged more than $3$ pounds?




\vspace{3cm}

### 4. One-sided alternative t-test 

Suppose we purchase $n=6$ random bags of chips marked with a net weight of $35.4$ grams. We then carefully weighing the contents in each bag to find the following weight data (in grams):

```{r echo=TRUE}
chips = c(35.5, 35.3, 35.1, 36.4, 35.4, 35.5)
```

a) Is there evidence that the mean weight of these chip bags is less than advertised? 

b) What would a Type II error mean in this situation? 


\vspace{3cm}

### 5.  Two-sided alternative t-test 

The National Perinatal Statistics Unit of the Sydney  Children’s Hospital reports that the mean birth weight of all  babies born in birth centers in Australia in a recent year was $3564$ grams—about $7.86$ pounds. A Missouri hospital reports that the average weight of $112$ babies born there last year was $7.68$ pounds, with a standard deviation of $1.31$ pounds. We want to see if U.S. babies weigh the same, on average, as Australian babies.

a) Why is this a one-sample test?

b) If we believe the Missouri babies fairly represent American newborns, is there any evidence that U.S. babies and Australian babies do not weigh the same at birth?



