---
title: "In-Class Worksheet"
author: "STAT011 with Prof Suzy"
date: "Week 3: Fitting a Linear Regression Model"
output:
  html_document: default
---
<br>
**Name:** _____________________________________

<br>
**Instructions:** Wildlife researchers are monitoring a Florida alligator population by taking areal photographs and attempting to estimate the weights of the gators based on the length of the gators in the images. The data set `Gators.csv` contains the variables `Length` and `Weight` for a sample of alligators who have been captured and studied. This data is shown in the scatterplot below. Import the `Gators` data set into either Excel or RStudio and then answer the following questions. 
<br>

```{r setup, echo=FALSE, include=FALSE}
require(knitr)
library(tidyverse)
knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings
```

```{r echo=FALSE, verbose=FALSE}

gators <- read.csv("https://raw.githubusercontent.com/ProfSuzy/Stat11/main/Data/Gators.csv")

ggplot(gators, aes(x=Weight.lbs., y=Length.in.)) +
  geom_point() + 
  labs(title = "Weight and Length of Captured Alligators") +
  xlab("Weight (lbs)") + ylab("Length (in)")

#lm(Weight.lbs.~Length.in., gators)$residuals %>% abs() %>% max
```
<br>

# 1. Choose which variable should be the response and justify this choice in 1-2 sentences.  

<br>
<br>
<br>
<br>

# 2. What are the slope and intercept of the line of best fit through this data? 

<br>
<br>
<br>
<br>

# 3.  Does the linear model seem to be a good fit for this data? If so, describe why. If not, what could be done to make a linear model more appropriate? 

<br>
<br>
<br>
<br>

# 4. The largest residual has a value of 165.62 Explain the meaning of this value in 1-2 sentences. 

<br>
<br>
<br>
<br>

# 5. If you were a wildlife researcher who needed to know the different weights of alligators, would you decide to use this method? Give a statistically informed justification of your answer.  

<br>
<br>
<br>
<br>