---
title: "In-Class Worksheet - Solutions"
author: "STAT011 with Prof Suzy"
date: "Week 4: Fitting a Linear Regression Model"
output:
  html_document: default
  pdf_document: default
---

**Instructions:** Wildlife researchers are monitoring a Florida alligator population by taking areal photographs and attempting to estimate the weights of the gators based on the length of the gators in the images. The data set `Gators.csv` contains the variables `Length` and `Weight` for a sample of alligators who have been captured and studied. This data is shown in the scatterplot below. Import the `Gators` data set into either Excel or RStudio and then answer the following questions. 
<br>

```{r setup, echo=FALSE, include=FALSE}
require(knitr)
library(tidyverse)
knitr::opts_chunk$set(message = FALSE) # include this if you don't want markdown to knit messages
knitr::opts_chunk$set(warning = FALSE) # include this if you don't want markdown to knit warnings
```

## Import data in Excel 

Copy and past the data from the following link: https://raw.githubusercontent.com/dr-suz/Stat11/main/Data/Gators.csv

## Import data in R 
```{r}
gators <- read.csv("https://raw.githubusercontent.com/dr-suz/Stat11/main/Data/Gators.csv")
```


## Scatter plot of the data 

```{r echo=FALSE, verbose=FALSE}
ggplot(gators, aes(x=Weight.lbs., y=Length.in.)) +
  geom_point() + 
  labs(title = "Weight and Length of Captured Alligators") +
  xlab("Weight (lbs)") + ylab("Length (in)")

#lm(Weight.lbs.~Length.in., gators)$residuals %>% abs() %>% max
```
<br>

# 1. Choose which variable should be the response and justify this choice in 1-2 sentences.  

\vspace{2cm}

# 2. What are the slope and intercept of the line of best fit through this data? What is the interpretation of the slope within this context? 

```{r}
slr <- lm(gators$Weight.lbs. ~ gators$Length.in.)
summary(slr)
```

# 3.  Does the linear model seem to be a good fit for this data? If so, describe why. If not, what could be done to make a linear model more appropriate? 

If we plot the data and the regression line, it becomes obvious that the relationship between length (in inches) and weight (in lbs) is non-linear. 

```{r}
plot(gators$Length.in., gators$Weight.lbs.)
abline(slr)
```

This is a sign that we should try transforming one or more of the variables to see if we can find a more linear relationship. For example, I'm going to transform the response variable by taking the natural logarithm (the function in R for this is `log()` ):

```{r}
slr_transformed <- lm(log(gators$Weight.lbs.) ~ gators$Length.in.)
summary(slr_transformed)
plot(gators$Length.in., log(gators$Weight.lbs.))
abline(slr_transformed)
```


This transformation worked! Our estimated regression equation for the transformed data is
$$\hat{ln(Weight)} = 1.34 + 0.035(length).$$ 
Now the interpretation of the slope is: for each additional inch in length (in the photographs), we expect the natural logarithm of the weight of the aligator to increase by $0.035$ lbs. We can un-do the transformation by raising $e$ to the power of the left hand side and the right hand side of our regression equation:
$$e^{\hat{ln(Weight)}} = \hat{Weight} = e^{1.34 + 0.035(length)} = e^{1.34} \times (e^{0.035})^{length}.$$
From simplifying this equation we can tell that the effect of an additional inche in length is actually a multiplicative increase in weight by $e^{0.035}=1.036$ lbs, on average. 

# 4. The largest residual has a value of 165.62. Explain the meaning of this value in 1-2 sentences. 

\vspace{3cm}

# 5. If you were a wildlife researcher who needed to know the different weights of alligators, would you decide to use this method? Give a statistically informed justification of your answer. (Hint: Report and interpret the $R^2$ value and/or the correlation coefficient.)

Even though the $R^2$ values is large, I would not use the original linear regression equation to estimate gator weight because the relationship between length (in inches) and weight (in lbs) is non-linear. 
