---
title: "Class 21 Group Worksheet"
subtitle: "Stat 11 with Prof. Suzy"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
```

Consider the data set in the `forcats` package called `gss_cat`. This is a sample of data from the General Social survey [General Social survey](https://gssdataexplorer.norc.org/)

First install the `forcats` package. 
```{r}
# install.packages("forcats") ## Uncomment this line and run it by itself once. Then comment it out again and knit your document. 
```

Then we can view the first few rows of the data set. 
```{r}
library(forcats)
head(gss_cat) 
```


We can create a contingency table with the following code: 

```{r}
gss_cat %>% group_by(marital, relig) %>%  ## specify which categorical variables you want to tabulate
            summarise(counts = n())  %>%  ## this specifies that we are creating a feature of our data set called "counts" and this is computed by counting up the number of observations using the function n()
            spread(relig, counts)    ##specify which counts you want to span the columns, this step is not necessary if you are creating a table of counts for a single categorical variable 
```

Note that you have to click the little black arrow in the upper corner of the table to see all the different levels associated with the variable  `relig`.

***
## 1. How many categorical variables are there in this data set and what are their names?  

[Write your answer here.]

## 2. How many levels are there for the categorical variable "relig" and what are their names?

[Write your answer here.]

## 3. Create a vector representing the different marital status counts for individuals who practice Hinduism. 

```{r}
#martical_counts <- c()
#martical_counts
```

## 4. Create a vector representing the counts of different religions represented for individuals who have divorced.  

```{r}
# religious_counts <- c()
# religious_counts
```


## 5. Complete the code below to conduct a Chi-squared test for homogeneity to determine if peoples' religions are distributed evenly among this sample. 

```{r}
#gss_cat %>% group_by(?) %>%   
#            summarise(counts = n()) 
#religions_counts <- c( )

#chisq.test(x = religions_counts)
```
