---
title: "Project Homework 4 Solutions"
author: "STAT011 with Prof Suzy"
output:
  pdf_document: default
  html_document: default
---



# Instructions

If you are analyzing this data in Excel you first need to [download the data set for HW 4](https://github.com/dr-suz/Stat11/blob/main/Data/gardasil_data.txt) from our Stat 11 Github Data page. Do this by right clicking on the link "View Raw" and save the link with the name `gardasil.csv`. 

If you are analyzing this data in RStudio, you will import the data with the following command
```{r, eval=TRUE}
gardasil <- read.delim(
        "https://raw.githubusercontent.com/dr-suz/Stat11/main/Data/gardasil_data.txt",
        sep="\t")
```
The data object is called `gardasil`. 



# Problem 1 

Consider the mean age of $11$ to $26$ year old female patients who visited the clinic because they were interested in getting the HPV vaccine. 

(a) Do do think that the two conditions necessary for the Central Limit Theorem to hold are reasonable in this example? Why/why not? 

(b) Regardless of your answer to (a), find an $85\%$ CI for the mean age of these patients. 

## Solution 1 

(a) To use the Central Limit Theorem we need to assume that the patients ages are all independent of one another and that this sample, although it is not random, is somehow representative of a larger population. This may not be unreasonable to assume. (This sample size is also less than $10\%$ the size of any such larger population.) Because our sample size is so large, even though the age of the patients is concentrated around 18 years, we can use the CLT to approximate the probabilities associated with the mean age of this sample. 

(b) An 85\% CI for the mean age of these patients is 18.385 years to 18.709 years. 

One could use the code:
```{r}
t.test(gardasil$Age, conf.level=0.85)
```
Or, equivalently, 
```{r}
n=1413 
xbar = mean(gardasil$Age)
SE = sd(gardasil$Age)/sqrt(n)
t_crit = qt((1-0.85)/2, df=n-1, lower.tail=FALSE)

xbar - (t_crit*SE)
xbar + (t_crit*SE)
```

### For graders

For full credit, answers must include a written statement noting the confidence interval within the context of the problem. 


# Problem 2 

Find a $90\%$ confidence interval for $p=$ the probability that an $11$ to $26$ year old female will complete all three required shots. Interpret this interval within the context of the problem. 

## Solution 2

We are 90\% confident that the proportion of 11-26 year old females who are interested in the HPV vaccine and actually complete the three required shots is between 31.17\% and 35.29\%.  

One could use the following code:
```{r}
table(gardasil$Completed)
prop.test(469, n=1413, conf.level = 0.9, correct=FALSE) 
```
Or, equivalently (except for small rounding error in the lower bound), 
```{r}
table(gardasil$Completed)
n = 1413
x = 469
p_hat = x/n
z_crit = qnorm((1-0.9)/2,0,1,lower.tail=FALSE)
SE = sqrt(p_hat*(1-p_hat)/n)
p_hat - (z_crit * SE)
p_hat + (z_crit * SE) 
```


### For graders

For complete credit, answers must state the resulting interval within the context of the problem. 


# Problem 3 

Determine how many **more** patients we would need to survey in order to find a $95\%$ confidence interval for the probability that a patient completes all three shots to within a margin of error of $0.1$? (Hint: First, figure out how many of the patients in this sample have completed all three shots. Then, continue to solve the sample size estimation problem.)



## Solution 3 

$\hat{p}=\frac{\text{number of patients who have completed all three shots}}{\text{number of patients}}$ 

**Version 1)** 

```{r}
table(gardasil$Completed)
n_completed = 469
n_total = 1413
  
ME = 0.1
z_star = qnorm((1-0.95)/2,0,1,lower.tail=TRUE)
p_hat = n_completed / n_total 

(p_hat*(1-p_hat))/(ME/z_star)^2
```
The required sample size is $n \geq$ `r round((p_hat*(1-p_hat))/(ME/z_star)^2,2)`, which is smaller than the sample we already have. Thus, we don't need any additional patients to produce the required CI. 

**Version 2)**  

This version gives a slightly more conservative sample size estimate using a $\hat{p}$ value set to $0.5$. 

```{r}
ME = 0.1
z_star = qnorm((1-0.95)/2,0,1,lower.tail=TRUE)
p_hat = 0.5 

(p_hat*(1-p_hat))/(ME/z_star)^2
```
The required sample size is $n \geq$ `r round((p_hat*(1-p_hat))/(ME/z_star)^2,2)`, which is smaller than the sample we already have. Thus, we don't need any additional patients to produce the required CI. 



### For graders: 

Please grade this problem for **completion** rather than for correctness.


# Problem 4 

Conduct a hypothesis test to determine if the average number of shots per patient is three or less than three. Choose your own confidence (and significance) level.

(a) State the null and alternative hypotheses. 

(b) Assess the necessary assumptions and conditions. 

(c) Regardless of your assessment in part (b), conduct the test in part (a) and state the p-value and your conclusion within the context of the problem. 

## Solution 4 

(a) Let $\mu =$ average number of shots completed by interested 11-26 year old patients. We are testing 
$$H_0: \mu = 3, \quad\text{vs}\quad H_A:\mu < 3.$$ 

(b) To use the Central Limit Theorem we need to assume that the patients ages are all independent of one another and that this sample, although it is not random, is somehow representative of a larger population. This may not be unreasonable to assume. (This sample size is also less than $10\%$ the size of any such larger population.) Because our sample size is so large, even though the age of the patients is concentrated around 18 years, we can use the CLT to approximate the probabilities associated with the mean age of this sample. 


(c) The p-value for this test is incredibly small (basically zero) therefore, at any level of significance, we reject the null in favor of the alternative. This means that based on these data, there is evidence suggesting that the average number of shots completed by this population is less than the three required. 

One could use the following code:
```{r}
t.test(gardasil$Shots, mu=3, alternative="less")
```
Or equivalently, 
```{r}
n=1413
TestStat = (mean(gardasil$Shots) - 3)/(sd(gardasil$Shots)/sqrt(n))
pt(TestStat, df=n-1, lower.tail = TRUE)
```


### For graders

Although the p-values differ slightly in the code above, they are both essentially zero. Please give credit for any answers to part (c) that state that the p-vale is approximately zero. 



# Problem 5 

Based on this survey, conduct a $0.01$ significance level hypothesis test of 
$$H_0: p=0.5 \quad \text{vs}\quad H_{A}: p< 0.5$$
where $p=$ the probability that an $11$ to $26$ year old female who is interested in protection against HPV does not have any medical assistance. State the p-value and your conclusion within the context of the problem. 

## Solution 5

One could use the following code: 
```{r}
table(gardasil$MedAssist)
prop.test(1138, n=1413, p=0.5, alternative="less", correct=FALSE)
```

Or equivalently, 
```{r}
table(gardasil$MedAssist)
n = 1413 
x = 1138

TestStat = ((x/n) - 0.5)/sqrt(.5*.5/n)

pnorm(TestStat, lower.tail=TRUE)
```

The p-value for this test is actually $1$ therefore, at any level of significance, we fail to reject the null hypothesis. This means that based on these data, there is no evidence to suggest that the proportion of patients without medical assistance is less than 50\%.  


### For graders 

For complete credit, answers must state the correct result and interpret it in the context of the problem. 
